---
title: "Practical-Machine-Learning-Project"
output: html_document
---
```{r}
library(caret)
library(randomForest)
```

 From the basic to creat the model which includes setting the working directory
 Loading the data into the console from the respective files
```{r}

setwd("E:/R-language/Practices/8part(Machine learning)/Project")

pmltrain <- read.csv("pml-training.csv",na.string=c("NA",""))

dim(pmltrain)
str(pmltrain)

```

##Data Cleaning
Here we can see that datahas 160 variable and most of them contains NA so Discarding the columns which have values NAs

```{r}

NAs <- apply(pmltrain, 2, function(x) { sum(is.na(x)) })

pmltrain_valid<-pmltrain[,which(NAs==0)]

dim(pmltrain_valid)
```
Also removed the colums "timestamp,x column,username,new window"" because they are not sensor values so the resulting data is as below
```{r}
removeIndex <- grep("timestamp|X|user_name|new_window", names(pmltrain_valid))
pmltrain_valid_rl <- pmltrain_valid[, -removeIndex]

dim(pmltrain_valid_rl)
```
we are left with only 54 variables now futher creating the subsets of data as train and test 

```{r}

Intrain <-createDataPartition(y=pmltrain_valid_rl$classe,p=0.2,list=FALSE)
train<-pmltrain_valid_rl[Intrain,]
test<-pmltrain_valid_rl[-Intrain,]
```

##cross validation
Applied cross validation using 10-fold cross validation
```{r}
#Configuration the train control for cross validation
cV <- trainControl(method="cv",number = 4)

#Fit the model using Random Forest
modFit<- train(train$classe~.,data=train,method="rf",trControl=cV
               ,prox=TRUE,allowParallel=TRUE)
#modFit
print(modFit)
print(modFit$finalModel)

```
##Predictions on the testing dataset

```{r}
#loading test data
pmltest <- read.csv("pml-testing.csv",na.string=c("NA",""))

#taking columns which are present in train data
test_rl<-pmltest[ ,which(names(pmltest) %in% names(train))]

#Run the prediction #Result
#apply random forest model to test set
pred<-predict(modFit,test_rl)
pred

# Utility function provided by the instructor
pml_write_files = function(x){
  n = length(x)
  for(i in 1:n){
    filename = paste0("problem_id_",i,".txt")
    write.table(x[i],file=filename,quote=FALSE,row.names=FALSE,col.names=FALSE)
  }
}

pml_write_files(pred)
```
We apply the randomForest model to the 20 given test set for the predictions. The results were all correct.


